# Buffering of stress reinstatement

The manipulation checks showed that that the stress effects of the first reminder were clearer, especially for feeling stressed. Therefore, the analysis of buffer effects focuses on the first reminder.

Here we investigate whether oxycodone reduces negative affective states induced by the 1st stress reminder (or *reinstatement*).

First we prepare the data:

```{r}
my_data.Q_buffer = 
  my_data.Q %>% 
  .[stage %in% c(5,6) & 
      item_name %in% c(negaffect_items,
                       posaffect_items,
                       embselfcn_items,
                       "stressed")] %>% 
  .[, stress_reminder := ifelse(stage %in% c(5,7) & state.condition == "stress","pre","post")] %>%
  .[, stress_reminder := factor(stress_reminder, levels = c("pre","post"))] %>% 
  .[, reminder := ifelse(stage %in% c(5,7),"pre","post")] %>% 
  .[, reminder := factor(reminder, levels = c("pre","post"))] %>% 
  .[, affect := ifelse(item_name %in% negaffect_items,
                       "negative",
                       ifelse(item_name %in% posaffect_items,
                              "positive","emb_selfc"))] %>% 
  .[item_name == "stressed", affect := "stressed"] %>% 
  coarsen(var = "response",
          bins = 20,
          range = c(0,100))
```

Here is a plot of the raw data.

```{r, fig.cap="Mean ratings and 95% confidence intervals for stress, embarrassment and self-conciousness pre and post 1st reminder by drug, state condition and sex.", fig.height=15}
my_data.Q_buffer_stats = 
  my_data.Q_buffer %>% 
  .[, .(response = mean(response),
        se = sd(response)/sqrt(.N)),
    by = .(Drug, reminder, item_name, Sex, state.condition)] %>% 
  .[, `:=`(lower = response-2*se,
           upper = response+2*se)] %>% 
  .[, grp := paste(Drug,state.condition)]

my_data.Q_buffer_stats %>% 
  ggplot(aes(x = reminder, y = response, color = Drug, lty = state.condition, 
             shape = state.condition, group = grp,
             ymin = lower, ymax = upper)) +
  geom_point(position = pd.2) + 
  geom_line(aes(group = grp), position = pd.2) +
  geom_linerange(position = pd.2) + 
  facet_grid(item_name~Sex, scales = "free_y") +
  scale_col_drug
```

```{r, fig.cap="Mean and 95% confidence interval for stress reminder effect, defned as the difference in the change in post-drug ratings from pre to post 1st reminder between the stressful and the control reminders. Stratified by drug condition (colors) and sex (panels). Note that these are calculated from the raw data.", fig.height=15, results='hide',fig.keep='all'}
pdata = 
  my_data.Q_buffer %>% 
  # average the pre stages BL1, BL2, BL3
  .[, .(response = mean(response)), 
    by = .(participant, Drug, state.condition, stage, Sex,item_name )] %>% 
  # calculate state induction effect as post-pre difference
  dcast(participant  + Drug + state.condition + Sex + item_name ~ stage, value.var = "response") %>%
  .[, induction_effect := `6` - `5`] %>% 
  # calculate stress effect as difference of post-pre difference in stress & control conditions
  dcast(participant + Drug + Sex + item_name ~ state.condition, value.var = "induction_effect") %>% 
  .[, stress_effect := stress - control]

pdata.stats = 
  seWithin(pdata, "stress_effect",
           withinvars = c("Drug","item_name"),
           betweenvars = c("Sex"),
           idvar = "participant") %>% 
  data.table() %>% 
  .[, stress_effect := as.numeric(stress_effect)] %>% 
  .[, `:=`(lower=stress_effect-ci,
           upper=stress_effect+ci)]

pdata.stats %>% 
  ggplot(aes(x = Drug, y = stress_effect, fill = Drug)) + 
  geom_bar(stat = "identity", alpha = .5) + 
  geom_errorbar(aes(ymin = lower, ymax = upper), width = .25, linewidth = 1) + 
  facet_grid(item_name~Sex) + 
  geom_quasirandom(data = pdata, aes(color = Drug), na.rm=TRUE) +
  scale_col_drug +
  coord_cartesian(ylim=c(-50,50))+ylab("Stress effect")
```

```{r, fig.cap="Distribution of the difference in stress reminder effect between drug conditions, stratified by sex. Note that these are calculated from the raw data"}
pdata %>% 
  .[, stress_effect := stress - control] %>% 
  # calculate drug x stress interaction effect as difference of stress effects between drug conditions
  dcast(participant + item_name + Sex ~ Drug, value.var = "stress_effect") %>% 
  .[, stress_x_drug := oxycodone - placebo] %>% 
  ggplot(aes(x = stress_x_drug, fill=Sex)) + 
  geom_histogram(alpha = .5, position = "identity", bins = 20) + 
  scale_col_sex + 
  facet_wrap(~item_name)+
  xlab("Stress*Drug interaction")
```

## Analysis approach


The general approach for the analyses below is to first obtain posterior distributions for the ratings in the 8 cells obtained by crossing:

-   stress vs control
-   pre vs post reminder
-   oxycodone vs placebo
-   men vs women

Then we can calculate the desired contrasts by calculating the pre to post change for the following 4 conditions from crossing stress with drug. 

- `pre_post_placebo_control = Drugplacebo:Stresscontrol:reminderpre - Drugplacebo:Stresscontrol:reminderpost` 
- `pre_post_placebo_stress = Drugplacebo:Stressstress:reminderpre - Drugplacebo:Stressstress:reminderpost` 
- `pre_post_oxycodone_control = Drugoxycodone:Stresscontrol:reminderpre - Drugoxycodone:Stresscontrol:reminderpost` 
- `pre_post_oxycodone_stress = Drugoxycodone:Stressstress:reminderpre - Drugoxycodone:Stressstress:reminderpost`

With those, we can calculate the stress reminder effects in the drug and placebo condition:

-   `stress_effect_placebo = pre_post_placebo_stress - pre_post_placebo_control`
-   `stress_effect_oxycodone = pre_post_oxycodone_stress - pre_post_oxycodone_control`

Finally, we can calculate whether the stress-reminder effect is smaller in the the oxycodone compared to the placebo condition:

-   `stress_x_oxycodone = stress_effect_oxycodone - stress_effect_placebo`

This description does not include sex effects, but the above contrasts can be calculated separately by sex.

*Note that given this specification of the contrasts, a buffer effect where post-reminder ratings of negative items like stress are lower under oxycodone compared to placebo will result in negative values for the stress_x_drug interaction effect.*

Here is the analysis model:

```{r class.source = 'fold-show'}
analysis_formula = 
  response.c ~ Sex:Drug:Stress:reminder + 
               (1|session) + 
               (1 + Drug:Stress:reminder | participant) 
```

## Buffering of stress

We focus on buffering of stress from the first reminder.

```{r est_buffer_stress1}
f = "fits/buffer/stressed.Rdata"
the_data = my_data.Q_buffer[item_name == "stressed"]
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r}
cat(paste0(length(unique(fit$data$participant)), " participants were included in this analysis."))
```

```{r, fig.cap = "Summary of fit indicators. R-hat should be smaller than 1.01 and the effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show a table with the complete sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```


```{r, fig.cap="Posterior predictive check of pre-post differences. Points are differences directly calculated from the data, intervals show differences calculated from model parameters."}
p = plot_buffer_effect(fit,the_data,"stressed")
p$ppc
```


The following figure shows stress-drug interactions (negative values indicate a buffering effect). The data support a hypothesis of no effect of oxycodone on the stress response to the reminder if the credible interval falls within a region of practical equivalence of +/- 5 points.

```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'stressed' ratings by state, drug and sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) on 'stressed' ratings with drug condition. Dashed vertical lines indicate a region of practical equivalence."}
((p$stress.plot + 
    ylab("stress effect on 'stressed' ratings.")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for 'stressed' ratings after reminders.") 
```


## Buffering of embarrassment

```{r est_buffer_emb}
f = "fits/buffer/emb.Rdata"
the_data = my_data.Q_buffer[item_name == "embarassed"]
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r}
cat(paste0(length(unique(fit$data$participant)), " participants were included in this analysis."))
```

```{r, fig.cap = "Summary of fit indicators. R-hat should be smaller than 1.01 and the effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show a table with the complete sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```


```{r, fig.cap="Posterior predictive check of pre-post differences. Points are differences directly calculated from the data, intervals show differences calculated from model parameters."}
p = plot_buffer_effect(fit,the_data,"embarassed")
p$ppc
```

```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'embarrassed' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) on 'embarrassed' ratings with drug condition. Dashed vertical lines indicate a region of practical equivalence."}
((p$stress.plot + 
    ylab("stress effect on 'embarrassed' ratings.")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for 'embarrassed' ratings after reminders.") 
```


## Buffering of self-conciousness

```{r est_buffer_selfcn}
f = "fits/buffer/selfcn.Rdata"
the_data = my_data.Q_buffer[item_name == "selfconscious"]
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r}
cat(paste0(length(unique(fit$data$participant)), " participants were included in this analysis."))
```

```{r, fig.cap = "Summary of fit indicators. R-hat should be smaller than 1.01 and the effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show a table with the complete sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```

```{r, fig.cap="Posterior predictive check of pre-post differences. Points are differences directly calculated from the data, intervals show differences calculated from model parameters."}
p = plot_buffer_effect(fit,the_data,"selfconscious")
p$ppc
```

```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'selfconscious' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) on 'selfconscious' ratings with drug condition. Dashed vertical lines indicate a region of practical equivalence."}
((p$stress.plot + 
    ylab("stress effect on 'selfconscious' ratings.")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for 'selfconscious' ratings after reminders.") 
```

## Buffering of stress, embarrassement & selfconciousness

Given that the response pattern for these three items is expected to be similar, we can obtain more stable estimates of a buffer effect by modeling all items jointly.

```{r est_buffer_embselfstr"}
f = "fits/buffer/embselfstress.Rdata"
the_data = my_data.Q_buffer[item_name %in% c(embselfcn_items,"stressed")]
analysis_formula = 
  response.c ~  Sex:Drug:Stress:reminder + (1|session) + (Drug:Stress:reminder| participant) + (1|item_name)
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r}
cat(paste0(length(unique(fit$data$participant)), " participants were included in this analysis."))
```

```{r, fig.cap = "Summary of fit indicators. R-hat should be smaller than 1.01 and the effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show a table with the complete sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```

```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'stressed', 'selfconscious' and 'embarrased' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) ratings with drug condition. Dashed vertical lines indicate a region of practical equivalence."}
p = plot_buffer_effect(fit,the_data,"stress_emb_selfc")
((p$stress.plot + 
    ylab("stress effect on 'stressed', 'emb.' & 'selfc.' ratings.")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for 'stressed', 'emb.' & 'selfc.' ratings after reminders.") 
```


## Buffering of negative mood

```{r est_buffer_negmood"}
f = "fits/buffer/negaffect.Rdata"
the_data = my_data.Q_buffer[item_name %in% negaffect_items]
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r}
cat(paste0(length(unique(fit$data$participant)), " participants were included in this analysis."))
```

```{r, fig.cap = "Summary of fit indicators. R-hat should be smaller than 1.01 and the effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show a table with the complete sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```


```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'negative mood' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) ratings with drug condition. Dashed vertical lines indicate a region of practical equivalence."}
p = plot_buffer_effect(fit,the_data,"negaffect")
((p$stress.plot + 
    ylab("stress effect on negative mood ratings")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for negative mood ratings after reminders.") 
```


## Buffering of positive mood


```{r est_buffer_posmood"}
f = "fits/buffer/posaffect.Rdata"
the_data = my_data.Q_buffer[item_name %in% posaffect_items]
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r}
cat(paste0(length(unique(fit$data$participant)), " participants were included in this analysis."))
```

```{r, fig.cap = "Summary of fit indicators. R-hat should be smaller than 1.01 and the effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show a table with the complete sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```

```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'positive mood' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) ratings with drug condition. Dashed vertical lines indicate a region of practical equivalence."}
p = plot_buffer_effect(fit,the_data,"posaffect")
((p$stress.plot + 
    ylab("stress effect on negative mood ratings")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for negative mood ratings after reminders.") 
```

