# Buffering of stress reinstatement

Here we analyse if Drug reduces stress-reinstatement induced negative emotions.

Prepare the data:

```{r}
my_data.Q_buffer = 
  my_data.Q %>% 
  .[stage %in% c(5,6) & 
      item_name %in% c(negaffect_items,
                       posaffect_items,
                       embselfcn_items,
                       "stressed")] %>% 
  .[, stress_reminder := ifelse(stage %in% c(5,7) & state.condition == "stress","pre","post")] %>%
  .[, stress_reminder := factor(stress_reminder, levels = c("pre","post"))] %>% 
  .[, reminder := ifelse(stage %in% c(5,7),"pre","post")] %>% 
  .[, reminder := factor(reminder, levels = c("pre","post"))] %>% 
  .[, affect := ifelse(item_name %in% negaffect_items,
                       "negative",
                       ifelse(item_name %in% posaffect_items,
                              "positive","emb_selfc"))] %>% 
  .[item_name == "stressed", affect := "stressed"] %>% 
  coarsen(var = "response",
          bins = 20,
          range = c(0,100))
```

Here is a plot of the raw data.

```{r, fig.cap="Mean ratings for stress, embarrassment and selfconciousness pre and post reminders by drug, state condition and sex.", fig.height=15}
my_data.Q_buffer_stats = 
  my_data.Q_buffer %>% 
  .[, .(response = mean(response),
        se = sd(response)/sqrt(.N)),
    by = .(Drug, reminder, item_name, Sex, state.condition)] %>% 
  .[, `:=`(lower = response-2*se,
           upper = response+2*se)] %>% 
  .[, grp := paste(Drug,state.condition)]

my_data.Q_buffer_stats %>% 
  ggplot(aes(x = reminder, y = response, color = Drug, lty = state.condition, 
             shape = state.condition, group = grp,
             ymin = lower, ymax = upper)) +
  geom_point(position = pd.2) + 
  geom_line(aes(group = grp), position = pd.2) +
  geom_linerange(position = pd.2) + 
  facet_grid(item_name~Sex, scales = "free_y") +
  scale_col_drug
```

```{r, fig.cap="Effect of stress reminders in the raw data measured after drug adminstration. Reminder effects are calculated as the change in stress ratings from pre to post stress induction and drug administration.", fig.height=15}
pdata = 
  my_data.Q_buffer %>% 
  # average the pre stages BL1, BL2, BL3
  .[, .(response = mean(response)), 
    by = .(participant, Drug, state.condition, stage, Sex,item_name )] %>% 
  # calculate state induction effect as post-pre difference
  dcast(participant  + Drug + state.condition + Sex + item_name ~ stage, value.var = "response") %>%
  .[, induction_effect := `6` - `5`] %>% 
  # calculate stress effect as difference of post-pre difference in stress & control conditions
  dcast(participant + Drug + Sex + item_name ~ state.condition, value.var = "induction_effect") %>% 
  .[, stress_effect := stress - control]

pdata.stats = 
  seWithin(pdata, "stress_effect",
           withinvars = c("Drug","item_name"),
           betweenvars = c("Sex"),
           idvar = "participant") %>% 
  data.table() %>% 
  .[, stress_effect := as.numeric(stress_effect)] %>% 
  .[, `:=`(lower=stress_effect-ci,
           upper=stress_effect+ci)]

pdata.stats %>% 
  ggplot(aes(x = Drug, y = stress_effect, fill = Drug)) + 
  geom_bar(stat = "identity", alpha = .5) + 
  geom_errorbar(aes(ymin = lower, ymax = upper), width = .25, linewidth = 1) + 
  facet_grid(item_name~Sex) + 
  geom_quasirandom(data = pdata, aes(color = Drug)) +
  scale_col_drug +
  ylim(-50,50)
```

```{r, fig.cap="Distribution of the interaction of stress effect (pre-post reminder) with drug condition (directly calculated from the data)."}
pdata %>% 
  .[, stress_effect := stress - control] %>% 
  # calculate drug x stress interaction effect as difference of stress effects between drug conditions
  dcast(participant + item_name + Sex ~ Drug, value.var = "stress_effect") %>% 
  .[, stress_x_drug := oxycodone - placebo] %>% 
  ggplot(aes(x = stress_x_drug, fill=Sex)) + 
  geom_histogram(alpha = .5, position = "identity", bins = 20) + 
  scale_col_sex + 
  facet_wrap(~item_name)
```

## Analysis approach


The analysis approach here is that we get posterior distributions for the ratings in the 8 cells obtained by crossing:

-   stress vs control
-   pre vs post reminder
-   drug vs placebo
-   men vs women

Then we can calculate the desired contrasts by first calculating the pre to post change for 4 conditions from stress x drug. 

- `pre_post_placebo_control = Drugplacebo:Stresscontrol:reminderpre - Drugplacebo:Stresscontrol:reminderpost` 
- `pre_post_placebo_stress = Drugplacebo:Stressstress:reminderpre - Drugplacebo:Stressstress:reminderpost` 
- `pre_post_drug_control = Drugdrug:Stresscontrol:reminderpre - Drugdug:Stresscontrol:reminderpost` 
- `pre_post_drug_stress = Drugdrug:Stressstress:reminderpre - Drugdrug:Stressstress:reminderpost`

With those, we can calculate the stress reminder effects in the drug and placebo condition

-   `stress_effect_placebo = pre_post_placebo_stress - pre_post_placebo_control`
-   `stress_effect_drug = pre_post_drug_stress - pre_post_drug_control`

And finally we can calculate if the stress-reminder effect is smaller in the the placebo compared to the placebo condition

-   `stress_x_drug = stress_effect_drug - stress_effect_placebo`

This description does not include sex effects, but these effects can be calculates separately by sex.

*Note that given this specification of the contrasts, a buffer effect where post-reminder ratings of negative items like stress are lower under oxycodone compared to placebo will result in negative values for the stress_x_drug interaction effect.*

Here is the analysis model:

```{r class.source = 'fold-show'}
analysis_formula = 
  response.c ~ Sex:Drug:Stress:reminder + 
               (1|session) + 
               (1 + Drug:Stress:reminder | participant) 
```

The analysis for manipulation checks showed that that the stress effects of the first reminder were clearer, especially for feeling stressed. Therefore, the analysis of buffer effects focuses on the first reminder.

## Buffering of stress

We focus on buffering of stress for the first reminder.

```{r est_buffer_stress1}
f = "fits/buffer/stressed.Rdata"
the_data = my_data.Q_buffer[item_name == "stressed"]
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r, fig.cap = "Summary of fit indicators. rhat should be smaller than 1.1 and effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show the table with the full sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```


```{r, fig.cap="Posterior predictive check of pre-post differences. Points are differences directly calculate from the data, intervals show differences calculate from model parameters."}
p = plot_buffer_effect(fit,the_data,"stressed")
p$ppc
```


The following figure shows stress-drug interactions (negative values indicate a buffering effect). The data support a hypothesis of no effect of oxycodone on stress response if the credible interval falls in a region of practical equivalence of +/- 5 points.

```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'stressed' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) on 'stressed' ratings with drug condition. Dashed vertical lines enclode a region of practical equivalence."}
((p$stress.plot + 
    ylab("stress effect on 'stressed' ratings.")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for 'stressed' ratings after reminders.") 
```


## Buffering of embarrassment

```{r est_buffer_emb}
f = "fits/buffer/emb.Rdata"
the_data = my_data.Q_buffer[item_name == "embarassed"]
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r, fig.cap = "Summary of fit indicators. rhat should be smaller than 1.1 and effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show the table with the full sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```


```{r, fig.cap="Posterior predictive check of pre-post differences. Points are differences directly calculate from the data, intervals show differences calculate from model parameters."}
p = plot_buffer_effect(fit,the_data,"embarassed")
p$ppc
```

```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'embarrassed' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) on 'embarrassed' ratings with drug condition. Dashed vertical lines enclode a region of practical equivalence."}
((p$stress.plot + 
    ylab("stress effect on 'embarrassed' ratings.")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for 'embarrassed' ratings after reminders.") 
```


## Buffering of selfconciousness

```{r est_buffer_selfcn}
f = "fits/buffer/selfcn.Rdata"
the_data = my_data.Q_buffer[item_name == "selfconscious"]
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r, fig.cap = "Summary of fit indicators. rhat should be smaller than 1.1 and effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show the table with the full sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```

```{r, fig.cap="Posterior predictive check of pre-post differences. Points are differences directly calculate from the data, intervals show differences calculate from model parameters."}
p = plot_buffer_effect(fit,the_data,"selfconscious")
p$ppc
```

```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'selfconscious' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) on 'selfconscious' ratings with drug condition. Dashed vertical lines enclode a region of practical equivalence."}
((p$stress.plot + 
    ylab("stress effect on 'selfconscious' ratings.")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for 'selfconscious' ratings after reminders.") 
```

## Buffering of stress, embarrassement & selfconciousness

Given that the data patterns for these three items are expected to be similar, we can obtain more stable estimates of a buffer effect by modeling all items jointly.

```{r est_buffer_embselfstr"}
f = "fits/buffer/embselfstress.Rdata"
the_data = my_data.Q_buffer[item_name %in% c(embselfcn_items,"stressed")]
analysis_formula = 
  response.c ~  Sex:Drug:Stress:reminder + (1|session) + (Drug:Stress:reminder| participant) + (1|item_name)
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r, fig.cap = "Summary of fit indicators. rhat should be smaller than 1.1 and effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show the table with the full sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```

```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'stressed', 'selfconscious' and 'embarrased' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) ratings with drug condition. Dashed vertical lines enclode a region of practical equivalence."}
p = plot_buffer_effect(fit,the_data,"stress_emb_selfc")
((p$stress.plot + 
    ylab("stress effect on 'stressed', 'emb.' & 'selfc.' ratings.")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for 'stressed', 'emb.' & 'selfc.' ratings after reminders.") 
```


## Buffering of negative mood

```{r est_buffer_negmood"}
f = "fits/buffer/negaffect.Rdata"
the_data = my_data.Q_buffer[item_name %in% negaffect_items]
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r, fig.cap = "Summary of fit indicators. rhat should be smaller than 1.1 and effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show the table with the full sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```


```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'negative mood', 'selfconscious' and 'embarrased' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) ratings with drug condition. Dashed vertical lines enclode a region of practical equivalence."}
p = plot_buffer_effect(fit,the_data,"negaffect")
((p$stress.plot + 
    ylab("stress effect on negative mood ratings")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for negative mood ratings after reminders.") 
```


## Buffering of positive mood


```{r est_buffer_posmood"}
f = "fits/buffer/posaffect.Rdata"
the_data = my_data.Q_buffer[item_name %in% posaffect_items]
if (file.exists(f)) {
  load(f)
} else {
  fit = fit_ordinal_model(
    the_data, analysis_formula, f, family = acat())
}
```

```{r, fig.cap = "Summary of fit indicators. rhat should be smaller than 1.1 and effect sample size (ess) should be larger than 500.", fig.height=3}
sample_diags = fit %>% model_summary()
sample_diags$plot
```

```{r results='asis', echo = FALSE}
if (rmd_output == "html") {
  cat("<details> \n <summary>Click to show the table with the full sample diagnostics.</summary>")
  cat(sample_diags$table)
  cat("</details> \n <br>")
}
```

```{r fig.width = 8, fig.cap = "A: Model-predicted effects of the first stress reminder on 'negative mood', 'selfconscious' and 'embarrased' ratings by state Drug and Sex. Points and lines are means and 66 & 95% Credible intervals. B: Posterior distribution of the interaction of stress effect (pre-post reminder) ratings with drug condition. Dashed vertical lines enclode a region of practical equivalence."}
p = plot_buffer_effect(fit,the_data,"posaffect")
((p$stress.plot + 
    ylab("stress effect on negative mood ratings")) | p$buffer.plot) + 
  plot_annotation(tag_levels = "A") + 
  plot_layout(widths = c(2,3))

p$buffer.stats %>% 
  my_flextable(caption = "Buffering effect of oxycodone for negative mood ratings after reminders.") 
```

